---
title: "Analysis on key factors affecting life satisfaction"
author: "Woolim Kim, Yena Joo, Guemin Kim"
date: "Oct 19, 2020"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
library(dplyr)
library(tidyverse)
data <- read_csv("gss.csv")
```

## Analysis on key factors affecting life satisfaction

# Woolim Kim, Yena Joo, Guemin Kim
# Oct 19, 2020

## Abstract

We use the 2017 General Social Survey(GSS) data to study the response of their _____ to ____. The target population includes all non-institutionalized persons 15 years of age and older, living in the 10 provinces of Canada. The frame population is everyone who is registered combining both landline and cellular with Statistics Canada’s address registers, and the sampled population is whoever was reached via telephone. The target populations were stratified into 27 strata by geographic areas—mostly CMA— and simple random sampling without replacement of records was performed in each stratum.  
We find….  
1. a strong ….  
2. no overall effects in ….  

Together, these results suggest that…….

## Introduction

Here is where you should give insight into the setting and introduce the goal of the analysis. Here you can introduce ideas and basic concepts regarding the study setting and the potential model. Again, this is the introduction, so you should be explaining the importance of the work that is ahead and hopefully build some suspense for the reader. You can also highlight what will be included in the subsequent sections.


## Data
Introduce the data, explain why it was selected. Make sure to comment on important features and highlight any potential drawbacks to the data.

** Guemin  
The data set we chose for this assignment contains responses of the General Social Survey conducted in 2017. The contents of the survey include some characteristics of diverse families in Canada, their socio-economic status, as well as other subjective information such as the respondent's life satisfaction and health conditions. 

Since we want to identify some key factors affecting one's life satisfaction as mentioned previously, the focus of our analysis will be "Health and subjective well-being". 

The reason for choosing the 2017 GSS is because it is the most recent^[As it is stated in the documentation of the GSS, one of the primary objectives of the General Social Survey is to monitor the well being of Canadians over time.
As a result, every survey conducted so far contain the responses related to the questions asking for the respondents' well being, and the 2017 GSS is the most recent survey with such reponses.] survey that includes  the "Health and subjective well being" concept.

Since the original data set contains too many variables that are not neccessary, we are going to clean the data set prior to analysis by removing them.  

```{r include = F, echo = F, warning = F, message = F}
# life satisfaction data with some potential factors
life_satisfaction_data <- data %>% 
  select(caseid, age, sex, feelings_life, self_rated_health, self_rated_mental_health, 
         income_family, average_hours_worked) %>% 
  mutate(self_rated_health = case_when(self_rated_health == "Excellent" ~ 5, 
                                       self_rated_health == "Very good" ~ 4, 
                                       self_rated_health == "Good" ~ 3, 
                                       self_rated_health == "Fair" ~ 2, 
                                       self_rated_health == "Poor" ~1, 
                                       self_rated_health == "Don't know" ~ 0)) %>%
  mutate(self_rated_mental_health = case_when(self_rated_mental_health == "Excellent" ~ 5, 
                                              self_rated_mental_health == "Very good" ~ 4, 
                                              self_rated_mental_health == "Good" ~ 3, 
                                              self_rated_mental_health == "Fair" ~ 2, 
                                              self_rated_mental_health == "Poor" ~1, 
                                              self_rated_mental_health == "Don't know" ~ 0)) %>%
  filter(self_rated_health != 0, self_rated_mental_health != 0, 
         average_hours_worked != "Don't know") %>%
  rename(ID = caseid, satisfaction_score = feelings_life, family_income = income_family) %>%
  mutate(self_rated_total_health = (self_rated_health + self_rated_mental_health)/2) #merged mental health and physical health in to one variable.
  
```

This is what our new data looks like:
```{r echo = F, warning = F, message = F}
head(life_satisfaction_data)
```

Our data contains 8 variables: ID, age, sex, satisfaction_score, self_rated_health, self_rated_mental_health, family_income, and average_hours_worked.  

Detailed descriptions of some variables:  
 * satisfaction_score indicates the life satisfaction score on a scale of 0(very dissatisfied) to 10(very satisfied).  
 * self_rated_health and self_rated_mental health are the physical and mental health ratings, respectively, on a scale of 1(poor) to 5(Excellent) given by the respondent.  
 * average_hours_worked indicates the average number of hours worked per week.  


Since we want to observe how the life satisfaction score is related to potential factors such as health or financial conditions, the response variable of our analysis will be satisfaction_score and the predictors will be the potential factors: self_rated_health, self_rated_mental_health, family_income, and average_hours_worked.
**



## Model

Introduce the selected model here. It is expected that you will use some mathematical notation here. If you do please ensure that all notation is explained. You may also want to discuss any special (hypothetical) cases of your model here, as well as any caveats.  

```{r}
# General Linear Regression Model
satisfaction_lm <- lm(satisfaction_score ~ as.factor(self_rated_health) 
                      + as.factor(self_rated_mental_health) 
                      + family_income 
                      + average_hours_worked, data = life_satisfaction_data)
summary(satisfaction_lm)

# Linear Regression Model 
# relationship between working hours and health condition
# (Response: self_rated_health, Predictor: average_hours_worked)
health_lm <- lm(self_rated_health ~ average_hours_worked, data = life_satisfaction_data)
summary(health_lm)

# Linear Regression Model (Predictor: physical and mental health ratings)
lm1 <- lm(satisfaction_score ~ as.factor(self_rated_health) 
          + as.factor(self_rated_mental_health), data = life_satisfaction_data)
summary(lm1)

# Linear Regression Model (Predictor: income of the family)
lm2 <- lm(satisfaction_score ~ family_income, data = life_satisfaction_data)
summary(lm2)

# Linear Regression Model (Predictor: )
lm3 <- lm(satisfaction_score ~ average_hours_worked, data = life_satisfaction_data)
summary(lm3)

immigrant_data <- data %>% select(caseid, age, sex, feelings_life, 
                                  age_immigration, landed_immigrant, citizenship_status) %>%
  filter(landed_immigrant != "Don't know", citizenship_status != "Don't know")
table(immigrant_data$citizenship_status)

# Logit Regression Model
# lm4 <- lm(feelings_life ~ as.factor(landed_immigrant), data = immigrant_data, family = "binomial")
# summary(lm4)

```

```{r}

life_satisfaction_data %>%
  ggplot(aes(x = as.factor(self_rated_total_health), y = (satisfaction_score))) + geom_boxplot() +
  ggtitle("How Mental and Physical Health Affects People's Satisfaction") +
  ylab("Satisfaction score") +
  xlab("Self rated total health")

```

## Results

Here you will include all results. This includes descriptive statistics, graphs, figures, tables, and model results. Please ensure that everything is well formatted and in a report style. You must also provide an explanation of the results in this section. You can overflow to an Appendix if needed. 

Please ensure that everything is well labelled. So if you have multiple histograms and plots, calling them Figure 1, 2, 3, etc. and referencing them as Figure 1, Figure 2, etc. in your report will be expected. The reader should not get lost in a sea of information. Make sure to have the results be clean, well formatted and digestible.

## Discussion

Here you will discuss conclusions drawn from the results and comment on how it relates to the original goal of the study (which was specified in the Introduction).

# Weaknesses

Here we discuss weaknesses of the study, data, analysis, etc. You can also discuss areas for improvement.

# Next Steps

Here you discuss subsequent work to be done after this report. This can include next steps in terms of statistical analysis (perhaps there is a more efficient algorithm available, or perhaps there is a caveat in the data that would allow for some new technique). Future steps should also be specified in terms of the study setting (eg. including a follow-up survey on something, or a subsequent study that would complement the conclusions of your report).


## References

1. GSS Data: 

2. Data Cleaning Code:




